\documentclass[11pt]{article} % use larger type; default would be 10pt


\title{Numerical Simulations Homework \texttt{\#} 2}
\author{Drew Coffin}
\date{February 1, 2017} 

\usepackage{cancel} 

\begin{document}
\maketitle

\section{Problem 1}

\subsection{Part a}

The given problem is 
\begin{equation} \label{eq:1}
\frac{\partial^2 T}{\partial x^2} + \frac{\partial^2 T}{\partial y^2} = 0, \qquad 0 < x < 1 \quad 0 < y < 1 
\end{equation}

We are given that $T(0, y) = T(1, y) = T(x,1) = 0$ and $T(x,0) = T_o$, i.e. a unit square with a value of 1 along the x-axis and $0$ along the other sides. Let us presume $T(x, y) = X(x)Y(y)$. Substitute into Eq. 1 and divide through by $T(x,y)$:

\[   \frac{1}{X} \frac{\partial^2 X}{ \partial x^2} +  \frac{1}{Y} \frac{\partial^2 Y}{\partial y^2} = 0 \]

\[   \frac{1}{X} \frac{\partial^2 X}{ \partial x^2} = -  \frac{1}{Y} \frac{\partial^2 Y}{\partial y^2} \]

This is satisfied by letting $X(x) = A \sin(n \pi x)$ with n as yet being undetermined. Meanwhile, $Y(y)$ has the form of a exponential decay; however no individual expression of the form $\sinh{y}$  will satisfy both conditions. Instead, we use a Fourier series of the form:

\[ T(x, y) = \sum_{n=1}^{\infty} A_n \sin(n \pi x) \sinh[n \pi \left( 1 - y)\right] \]

and $A_n$ is determined by
\[ A_n = \frac{2 T_o}{ \sinh(n \pi)} \int_{0}^{1} \sin(n \pi x) dx\]

\[ A_n = \frac{4 T_o}{n \pi  \sinh(n \pi)}\]
From the shape of the graph, only the odd $A_n$ terms contribute.

Now, we plot.

\subsection{Part c}
The given problem is 
\begin{equation} \label{eq:2}
\frac{\partial^2 T}{\partial x^2} + \frac{\partial^2 T}{\partial y^2} = 0, \qquad 0 < x < 1 \quad 0 < y < 1 
\end{equation}

We are given that $T(0, y) = T(1, y)  = 0$ and $T(x,0) = T(x,1) = T_o$, i.e. a unit square with a value of 1 along bothsides parallel to the x-axis and $0$ along the sides parallel to the y-axis. The solution follows precisely above, except we will use a $cosh(y)$ expression shifted to center on $y = 1/2$, to give us the symmetric "valley" shape.

I.e.

\[ T(x, y) = \sum_{n=1}^{\infty} B_n \sin(n \pi x)  \cosh[n \pi \left(y - 1/2)\right] \] 
with $B_n$ being \[ B_n = \frac{4 T_o}{n \pi  \cosh(n \pi / 2)}\]
Again, we only retain odd $B_n$ terms.

\section{Problem 2}

\subsection{Part a}

The given equation is $\partial^2 u/\partial x \partial t = 0$ and we wish to know what type of second-order PDE it is. The type is determined by the discriminant of the coefficients of the second-order derivatives of our function $u(x,y)$: 

\[ A \frac{\partial^2 u}{ \partial x^2} + B \frac{\partial^2 u}{ \partial x \partial y} + C \frac{\partial^2 u}{ \partial y^2} + D(x,y) = 0 \]

If $B^2 - 4 A C > 0$, our equation is hyperbolic.

If $B^2 - 4 A C = 0$, our equation is parabolic.

If $B^2 - 4 A C < 0$, our equation is elliptic.

$B^2 - 4 A C $ in our case will be $1^2 - 4 \cdot 0 \cdot 0  = 1 > 0 $ so our equation is hyperbolic.

\subsection{Part b}

The given system of equations is ${\partial v}/{\partial x} = 0$ and ${\partial u}/{\partial t} - v= 0$.

If we take a spatial derivative of the second equation, we achieve:

\[ \frac {\partial} {\partial x} \left( \frac {\partial u}{\partial t} - v \right) = 0\]
\[ \frac {\partial^2 u}{\partial x \partial t} - \cancelto{0}{\frac {\partial v}{\partial x}} = 0\]
\[  \frac {\partial^2 u}{\partial x \partial t} = 0\]
as above, hence this is a hyperbolic equation.

For finding characteristics, we wish to find total time derivatives. 

\[ \frac {\partial v}{\partial x} = 0 \]

\[\frac {\partial}{\partial x} \frac {d x}{dt}= 0 \]

\[\frac {d}{dt} \frac {\partial x}{\partial x} = 0 \]

We have lost our x-dependence, thus we know we have no information about our equation based on the domain. Thus, only our coordinates, and thus their axes, give us characteristics.

\end{document}
